---
title: "README"
author: "Bob Rudis"
date: February 11, 2015
output:
  md_document:
    variant: markdown_github
---

streamgraph is an htmlwidget for making streamgraphs. Planned support for `xts` objects and also fixing the tooltip display (doesn't work with the [sample Rmd](http://rpubs.com/hrbrmstr/streamgraph_01))

The following functions are implemented:
- `sg_axis_x`	Modify streamgraph x axis formatting
- `sg_axis_y`	Modify streamgraph y axis formatting
- `sg_colors`	Modify streamgraph colors
- `streamgraph`	Create a new streamgraph

### News

- Version 0.1 released

### Installation

```{r eval=FALSE}
devtools::install_github("hrbrmstr/streamgraph")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
options(width=120)
```

### Usage

```{r eval=FALSE}
library(streamgraph)

# current verison
packageVersion("streamgraph")

library(dplyr)

ggplot2::movies %>%
  select(year, Action, Animation, Comedy, Drama, Documentary, Romance, Short) %>%
  tidyr::gather(genre, value, -year) %>%
  group_by(year, genre) %>%
  tally() %>%
  ungroup %>%
  mutate(year=as.Date(sprintf("%d-01-01", year))) -> dat

streamgraph(dat, "genre", "n", "year", interactive=TRUE) %>%
  sg_axis_x(20, "year", "%Y") %>%
  sg_colors("PuOr")

```

### Test Results

```{r}
library(streamgraph)
library(testthat)

date()

test_dir("tests/")
```

