---
title: "README"
author: "Bob Rudis"
date: February 12, 2015
output:
  md_document:
    variant: markdown_github
---

streamgraph is an htmlwidget for making streamgraphs. Planned support for `xts` objects.

[Sample Rmd](http://rpubs.com/hrbrmstr/streamgraph_01)

A streamgraph (or "stream graph") is a type of stacked area graph 
which is displaced around a central axis, resulting in a flowing, 
organic shape. Streamgraphs were developed by Lee Byron and popularized 
by their use in a February 2008 New York Times article on movie
box office revenues. ([Wikipedia](http://en.wikipedia.org/wiki/Streamgraph))

The following functions are implemented:

- `streamgraph`	: Create a new streamgraph
- `sg_axis_x`	: Modify streamgraph x axis formatting
- `sg_axis_y`	: Modify streamgraph y axis formatting
- `sg_colors`	: Modify streamgraph colors

### News

- Version `0.1` released
- Version `0.2` released - working SVG tooltips; general code cleanup

### Installation

```{r eval=FALSE}
devtools::install_github("hrbrmstr/streamgraph")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
options(width=120)
```

### Usage

```{r eval=FALSE}
library(streamgraph)

# current verison
packageVersion("streamgraph")

library(dplyr)

ggplot2::movies %>%
  select(year, Action, Animation, Comedy, Drama, Documentary, Romance, Short) %>%
  tidyr::gather(genre, value, -year) %>%
  group_by(year, genre) %>%
  tally(wt=vale) %>%
  ungroup %>%
  mutate(year=as.Date(sprintf("%d-01-01", year))) -> dat

streamgraph(dat, "genre", "n", "year", interactive=TRUE) %>%
  sg_axis_x(20, "year", "%Y") %>%
  sg_colors("PuOr")

```

### Test Results

```{r}
library(streamgraph)
library(testthat)

date()

test_dir("tests/")
```

